Manuscript Draft\\ \vspace{2cm} What is long-distance dispersal? And a taxonomy of dispersal events

### 2016-06-28 Manuscript/MS- What is LDD.tex
Minor edits.

The movement of pollen and seeds by animals and its consequences have intrigued population geneticists and field ecologists since the infancy of both research disciplines. Each has generated an impressive body of theoretical and empirical research through the past decades, yet advances have long been co-existing in ‘parallel worlds’ and the great synergistic potential of population genetics and demography for the study of plant dispersal by animals remains little explored. Knowledge gaps still having the imprint of this conceptual disconnection include the idea of long distance dispersal, and the paradoxes of forest fragmentation effects on genetic diversity \citep{Kramer:2008kg}, survival and persistence of relict tree species \citep{Hampe:2011bv}, rapid post-Pleistocene recolonization of vast continental areas in response to climate modification \citep{Clark:1998aa,Clark:1998vi}, among other persisting issues. This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD). LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance. Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications [REF]. But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}. In particular, LDD events remain difficult to assess, both technically- with serious methodological problems for its reliable estimation- and conceptually. Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.

@@

No explicit definition of what constitutes an LDD event exists. Previous approaches \citep[e.g., ][]{Nathan:2006aa,Schurr2009long} include both absolute and proportional definitions to characterize LDD events. This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold \citep{Nathan:2008is}. The exact proportional or absolute thresholds selected remain arbitrary, as no reference spatial frame is provided within the definition of LDD. This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb}. For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}. Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb}.

We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of its frequency and extent could be comparable across different study systems. We argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses. These elements can be overlaid on previous definitions based on absolute and proportional characterizations of LDD. We start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds. We then use a case study to assess differential contributions of animal frugivores performing LDD.

@@

A second consideration in terms of spatial boundaries, with effects on dispersal patterns, is the genetic neighborhood area $N_b$, i.e., the spatial extent including a subset of panmictic individuals within a population \citep{Wright:1943aa,Wright:1946aa}. Thus, the $N_b$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all the individuals. However, most populations and stands of long-lived trees show highly aggregated and clumped distributions \citep{Seidler:2006hx}, where relatively long distances may separate groups of individuals within the same population. In these cases we might expect $N_b$ area to be substantially smaller than the total population area. Therefore, at least four possible scenarios exist with distinct implications in terms of consequences for dispersal (Table 1). In the case of dispersal events not extending beyond the geographic limits of the population or reference area, actual LDD events may involve dissemination beyond a reduced neighborhood area that is smaller than the geographic extent of the population, originating local long-distance ($LDD_{loc}$) dispersal events (Table 1, Fig. 2A). Actual short-distance dispersal would then involve those situations where the propagule is disseminated within \textit{both} the population limits and the genetic neighborhood boundary ($SDD_{loc}$). Along a similar reasoning, dispersal events outside the population limits will not necessarily convey LDD (Table 1, Fig. 2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa}, generating LDD events within the genetic neighborhood ($LDD_{neigh}$). Finally, strict-sense LDD events would involve dissemination outside \textit{both} the population limits and the genetic neighborhood boundary ($LDD_{ss}$) (Table 1, Fig. 2A).

@@

To a large extent, short-distance dispersal events (strict-sense, $SDD_{loc}$ events) are contributed by small- and medium-sized (\textit{Turdus}) frugivorous birds (Table 2). Given the relatively reduced $N^b_e$ area of \textit{P. mahaleb} (Suppl. Mat. Table S1), $< 1km^2$, well below the extent of the local study population \citep{Garcia:2007he,Garcia:2005fu},  we cannot estimate $LDD_{neigh}$ events (Table 2), as all LDD events outside the reference population occur, by definition, outside the $N^b_e$ area. Larger frugivores such as corvids and the pigeon \textit{Columba palumbus} contribute most LDD events, and most immigrant seeds potentially dispersed from other populations. Notably, strict-sense long-distance dispersal ($LDD_{ss}$) appears consistently associated with large-bodied frugivores (Table 2), most likely associated with a greater frequency of movements outside the local population (Fig. 4).

>>>>

The movement of pollen and seeds by animals and its consequences have intrigued population geneticists and field ecologists since the infancy of both research disciplines. Each has generated an impressive body of theoretical and empirical research through the past decades, yet advances have long been co-existing in ‘parallel worlds’ and the great synergistic potential of population genetics and demography for the study of plant dispersal by animals remains little explored. Knowledge gaps still having the imprint of this conceptual disconnection include the idea of long distance dispersal, and the paradoxes of forest fragmentation effects on genetic diversity \citep{Kramer:2008kg}, survival and persistence of relict tree species \citep{Hampe:2011bv}, rapid post-Pleistocene recolonization of vast continental areas in response to climate modification \citep{Clark:1998aa,Clark:1998vi}, among other persisting issues. This conceptual isolation has been exacerbated by technical difficulties for the robust characterization of dispersal events, especially those involving movement over long-distances (long-distance dispersal, LDD). LDD is a characteristically extreme event of propagule movement in any plant population, typically occurring with an extremely low probability but potentially reaching an extremely long distance. Some progress has recently been made through the fast-paced implementation of molecular tools in ecological research labs and the availability of cutting-edge technology for biotelemetry applications. But much of the population geneticist and ecologist communities remains unaware of the state of the art in each other and likely under-appreciates their potential to validate and enrich dispersal studies \citep{Jones:2008il}. In particular, LDD events remain difficult to assess, both technically- with serious methodological problems for its reliable estimation- and conceptually. Our aim here is to review the LDD concept with a specific emphasis on dispersal of plant propagules (seeds and pollen), providing an extended definition that might be helpful in the robust quantification of LDD events.

@@

No explicit definition of what constitutes an LDD event exists. Previous approaches \citep[e.g., ][]{Nathan:2006aa,Schurr2009long} include both absolute and proportional definitions to characterize LDD events. This means providing information about the absolute distances moved by a given percentile of the events and/or providing data on the proportion of events exceeding a given distance threshold \citep{Nathan:2008is}. The exact proportional or absolute thresholds selected remain arbitrary, as no reference spatial frame is provided within the definition of LDD. This leaves the consideration of LDD as an extreme form of context-dependent phenomenon, strongly dependent upon the scale of the biological process studied \citep{Kinlan:2005fb} and of the specific organism considered. For example, \cite{Kinlan:2005fb} used a spatial reference frame to characterize LDD events of marine organisms, where sedentary adults and larvae differ enormously in the spatial scales of their dispersal \citep{DAloia:2013fc}. Therefore, any measure of extent and reach of LDD events requires reference to an explicit spatial frame or "local" scale \citep{Kinlan:2005fb}.

We aim at providing a general framework for the quantitative analysis of LDD events so that estimates of their frequency and extent could be comparable across different study systems. We argue that both demographic and genetic elements are needed for this framework, most likely requiring a combination of field-based movement data and genetic analyses. These elements can be overlaid on previous definitions based on absolute and proportional characterizations of LDD. We start with a definition of LDD events within a spatially-explicit mechanistic framework allowing an unambiguous meaning for setting long-distance thresholds. We then use a case study to assess differential contributions of animal frugivores performing LDD.

@@

A second consideration in terms of spatial boundaries, with effects on dispersal patterns, is the genetic neighborhood area $N^b_e$, i.e., the spatial extent including a subset of panmictic individuals within a population \citep{Wright:1943aa,Wright:1946aa}. Thus, the $N^b_e$ area can be equal to the whole extent of the population whenever the population is unstructured and there is evidence for random mating events among all the individuals. However, most populations and stands of long-lived trees show highly aggregated and clumped distributions \citep{Seidler:2006hx}, where relatively long distances may separate groups of individuals within the same population. In these cases we might expect $N^b_e$ area to be substantially smaller than the total population area. Therefore, at least four possible scenarios exist with distinct implications in terms of consequences for dispersal (Table 1). In the case of dispersal events not extending beyond the geographic limits of the population or reference area, actual LDD events may involve dissemination beyond a reduced neighborhood area that is smaller than the geographic extent of the population, originating local long-distance ($LDD_{loc}$) dispersal events (Table 1, Fig. 2A). Actual short-distance dispersal would then involve those situations where the propagule is disseminated within \textit{both} the population limits and the genetic neighborhood boundary ($SDD_{loc}$). Along a similar reasoning, dispersal events outside the population limits will not necessarily convey LDD (Table 1, Fig. 2B): this is expected in cases where the genetic neighborhoods are extensive, going beyond the geographic limits of local populations, as in fig trees \citep{Nason:1998aa}, generating LDD events within the genetic neighborhood ($LDD_{neigh}$). Finally, strict-sense LDD events would involve dissemination outside \textit{both} the population limits and the genetic neighborhood boundary ($LDD_{ss}$) (Table 1, Fig. 2A).

@@

To a large extent, short-distance dispersal events (strict-sense, $SDD_{loc}$ events) are contributed by small- and medium-sized (\textit{Turdus}) frugivorous birds (Table 2). Given the relatively reduced $N^b_e$ area of \textit{P. mahaleb} (Suppl. Mat. Table S1), $< 1km^2$, well below the extent of the local study population \citep{Garcia:2007he,Garcia:2005fu},  we cannot estimate $LDD_{neigh}$ events (Table 2), as all LDD events outside the reference population occur, by definition, outside the $N^b_e$ area. Larger frugivores such as corvids and the pigeon \textit{Columba palumbus} contribute most LDD events, and most immigrant seeds potentially dispersed from other populations (Fig. S2). Notably, strict-sense long-distance dispersal ($LDD_{ss}$) appears consistently associated with large-bodied frugivores (Table 2), most likely associated with a greater frequency of movements outside the local population (Fig. 4).

Empirically mapping of dispersal events for either pollen or seed disseminated by animals may result in a complex pattern of different combinations of dispersal events (Fig. S1), as animal movements are overlaid onto plant populations occupying complex landscapes, resulting in different types of SDD and LDD events.
